---
title: "Project 2"
author: "Mina Kanaani"
date: "`r Sys.Date()`"
output: word_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

### Classification Project

This is the Report on the first R and Statistical Analysis's project on Classification and machine learning.

In this report I aim to predict the chance of occurring Stroke on hospital patients based on some important features such as their age,heart disease, hypertension(high blood pressure) ,glucose leve they have, etc.

Based on their characteristics, I used two methods : None Parametric methods such as KNN and Regression Tree and Parametric methods such as Logistic Regression and Discrimination models.

At the end of each method, I tested some important criteria based on problem's characteristic to examine the prediction to determine whether they showed us good results or not.

### Why Stroke?

This report is an analysis of a data set of hospital patients for the purpose of assigning risk factors to strokes and making recommendations to help prevent this event. Strokes are the second leading cause of death in the world according to the World Health Organization and are responsible for 11% of total deaths. Every year 15 million people suffer a stroke, 5 million of which pass away and another 5 million of which are left with a permanent disability (WHO, 2021). Many of these strokes are preventable through healthy habit forming and monitoring those at the highest risk can have a significant improvement in outcomes. For these reasons, it is an area that requires further study to prevent this event from impacting more lives than necessary. The following report focuses on determining risk factors for strokes and makes recommendations on how to prevent them.

## Flowchart

```{r}
library(DiagrammeR)

DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = TB, label = 'Project Flowchart',labelloc= ta]
node [shape = rectangle, style = filled, fillcolor = lightpink]

'Calling libraries' ->'Cleansing'-> 'EDA analysis' -> 'Processing data' -> 'None Parametric'
'Processing data' -> 'Parametric'
'Parametric'-> 'Logistic Regression'
'Parametric'-> 'Discrimination models'
'None Parametric' -> 'Regression Tree model'
'None Parametric' -> 'KNN model'
'KNN model'-> 'Visualization and testing'
'Regression Tree model'-> 'Visualization and testing'
'Logistic Regression'-> 'Visualization and testing'
'Discrimination models'-> 'Visualization and testing' 
'Visualization and testing'->'fitting and testing Candid model'->'Over Smapling and Under Sampling'
}")
```

## Recalling Libraries and Data

First we recall the libraries we need in this project.

```{r}
#needed libraries
library(readxl)
library(tibble)   # data frame printing
library(dplyr)      # data manipulation
library(caret)     # fitting knn
library(rpart)    # fitting trees
library(rpart.plot) # plotting trees
library(knitr)
library(Hmisc)
library(cowplot)
library(ggplot2)
library(MASS)
library(tidyverse)
library(knitr)
#library(kableExtra)
library(ROSE) 
#specific
library(e1071)
library(nnet)
library(pROC)
library(ellipse)
library(klaR)
#library(WVPlots)
#EDA needed library
library(ggplot2)
```

Then we read the excel file attached to this file to be able to access the data .

```{r}
#reading the file
heart_data=read_excel("healthcare-dataset-stroke-data.xlsx")
kable(head(heart_data,5))
```

## Cleansing

In this part I aimed to clean the data from discrepancies and nulls. I divided the cleaning part into 3 parts.

First part is cleaning the BMI features:

### BMI

```{r}
#first:BMI
# Check data type
class(heart_data$bmi)
# Convert BMI to numeric
heart_data$bmi=as.numeric(heart_data$bmi)
# Check data type
class(heart_data$bmi)
#View Summary Statistics for data
summary(heart_data$bmi)
#Replace N/A's in BMI column with mean
heart_data$bmi[is.na(heart_data$bmi)] =mean(heart_data$bmi,na.rm=TRUE)
sum(is.na(heart_data$bmi))
#View New Summary Statistics for data
summary(heart_data$bmi)
```

As It was shown in chunk above, there were 201 NA records in BMI features which were replaced with mean of BMI column.

part 2 of cleaning is for Gender feature:

### Gender

```{r echo=TRUE}
#second : Gender
table(heart_data$gender)
#we replace the "Other" category with "Female" since it's the majority of records.
heart_data$gender=ifelse(heart_data$gender == "Other", "Female", heart_data$gender)
kable(table(heart_data$gender))
```

As we can compare the two table, the Gender had a third category consist of "other' , since the female form the majority of data, we add the "other" to female category.

part 3 and last part of cleaning is for Smoking status.

### Smoking Status

Since there is an "unknown" category, we put the data in that column in other categories based on their probability. Calculate the probability of formerly smoker, current smokers and non-smokers given that there's only this three categories in the smoking_status column

```{r}
prob.FS = 885 / (885 + 1892 + 789)
prob.NS = 1892 / (885 + 1892 + 789)
prob.S =789 / (885 + 1892 + 789)
#Duplicate the data
data1=heart_data
#Replacing 'Unknown' in smoking_status by the other 3 variables according to their weightage
data1$rand <- runif(nrow(data1))
data1 = data1%>%mutate(Probability = ifelse(rand <= prob.FS, "formerly smoked", ifelse(rand <= (prob.FS+prob.NS), "never smoked", ifelse(rand <= 1, "smokes", "Check"))))
data1 = data1%>%mutate(smoking.status = ifelse(smoking_status == "Unknown", Probability, smoking_status))
table(data1$smoking.status)
#Remove columns that are not needed
heart_data=subset(data1, select = -c(rand,Probability,smoking_status))
# revise the column name of smoking status
colnames(heart_data)[12] = "smoking_status"
kable(head(heart_data,3))
```

## EDA Visualization

In this part, we use EDA analysis tools to visualize our features and their binary correlation with charges to determine which features have significant effect and which are not of much importance.

### Bar Chart

The first tool we use is Bar charts to illustrates the relationship between Quality features such as stroke, heart disease,sex,residence,smoke,work type and hypertension.

**1.**

```{r}
Yes=subset(heart_data, stroke == '1')
No=subset(heart_data, stroke == '0')
strokecount=as.data.frame(table(heart_data$stroke))
ggplot(strokecount, aes(x = Var1, y = Freq, fill = Var1)) +
  geom_bar(stat = "identity") + theme(legend.position="none") +
  geom_text(aes(label = Freq), vjust = 1.5) +
  labs(title="Stroke Status of Patients",x ="Stroke", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5))
```

As it is illustrated , the majority of record have not experienced stroke , only 249 people did so.

In conclusion we can see that we have an imbalance data.

**2.**

```{r}
#Create hypertension counts table

hypercount=as.data.frame(table(heart_data$hypertension,heart_data$stroke))
# Replace headers
colnames(hypercount)[1] ='Hypertension'
colnames(hypercount)[2] = 'Stroke'
hypercount$Hypertension =ifelse(hypercount$Hypertension == 0, "No", 'Yes')
hypercount$Stroke =ifelse(hypercount$Stroke == 0, "No", 'Yes')
# Bar Chart of Hypertension : No vs. Yes     
ggplot(hypercount, aes(x = Hypertension, y = Freq, fill = Stroke)) +
  geom_bar(stat = "identity")+
  geom_text(aes(label = Freq), vjust = 0) +
  labs(title="Hypertension Status of Patients",x ="Hypertension", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5))
```

We can have conclusion that people with blood pressure less than normal form the majority of our data, how ever the gap between the people who has hypertension and those who have not is slightly less than the plot with stroke patients. we can see that 15% of people with hypertension have experienced stoke and only 4 % of people with no hypertension had an stoke before. we can conclude the "hypertension" is one of important features effecting the stroke chance.

**3.**

```{r}
# Create heart disease counts table
heartcount =as.data.frame(table(heart_data$heart_disease, heart_data$stroke))
# Replace headers
colnames(heartcount)[1] <- 'Heart_Disease'
colnames(heartcount)[2] <- 'Stroke'
# Replace num to char
heartcount$Heart_Disease <- ifelse(heartcount$Heart_Disease== 0, "No", 'Yes')
heartcount$Stroke <- ifelse(heartcount$Stroke == 0, "No", 'Yes')

# Bar Chart of Heart Disease : No vs. Yes     
ggplot(heartcount, aes(x = Heart_Disease, y = Freq, fill = Stroke)) +
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Freq), vjust = 1) +
  labs(title="Heart Disease Status of Patients",x ="Heart Disease", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5))
```

Again as it can be seen, the vast number of record do not have heart disease , 20 % of people who Have Heart disease have had stroke and only 4 % of people without heart disease experienced stroke, which illustrates that Heart disease have effect on number of stoke Heart disease have a relatively important effect on chance of stroke.

**4.**

```{r}
# Create gender counts table
gendercount=as.data.frame(table(heart_data$gender,heart_data$stroke))
colnames(gendercount)[2] ='Stroke'
# Bar Chart of Gender   
ggplot(gendercount, aes(x = Var1, y = Freq, fill = Stroke)) +
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Freq), vjust = 1.5) +
  labs(title="Gender of Patients",x ="Gender", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5))
```

We can see the number of patients with stroke experience in male is slightly less than females which is logical due to more number of females in general, so we can conclude the gender has no significant effect on stroke.

**5.**

```{r}
# Create work type counts table
workcount = as.data.frame(table(heart_data$work_type))
# Bar Chart of Patient Work Type   
ggplot(workcount, aes(x = Var1, y = Freq, fill = Var1)) +
  geom_bar(stat = "identity") +  theme(legend.position="none") +
  geom_text(aes(label = Freq), vjust = 0) +
  labs(title="Patient Work Type",x ="Work Type", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5))
```

The majority of data, had private work type, the least type is 22 people who have never worked.There are approximately even amounts of patients that are working government jobs, are self-employed, and are children.

**6.**

```{r}
#effect of work type on stroke
work_stroke_count = as.data.frame(table(heart_data$work_type,heart_data$stroke))
# Bar Chart of Patient Work Type   
ggplot(work_stroke_count, aes(x = Var1, y = Freq, fill = Var2)) +
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Freq), vjust = 0) +
  labs(title="Patient Work Type",x ="Work Type", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5))
```

As the chart illustrates, the most patients with stoke experience, have the Private type of work, we can see 2 patients being under age had experience of stoke as well, not so much of a significant effect on stroke chance.

**7.**

```{r}
# Create ever married counts table
marriedcount <- as.data.frame(table(heart_data$ever_married,heart_data$stroke))
colnames(marriedcount)[2] <- 'Stroke'
# Bar Chart of Patients Who Have Been Married  
ggplot(marriedcount, aes(x = Var1, y = Freq, fill = Stroke)) +
  geom_bar(stat = "identity")  +
  geom_text(aes(label = Freq), vjust = 1) +
  labs(title="Bar Chart of Patients Who Have Been Married",x ="Have the Patient been Married", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5))
```

The majority of record have been married before.Roughly double the amount of patients have been married before than those who have not.as shown, almost 7% of married people experienced stroke whereas only 1% of single people experienced the stroke, so we can conclude there is a slightly effect of married on stroke chance.

**8.**

```{r}
# Create residence type counts table
rescount=as.data.frame(table(heart_data$Residence_type,heart_data$stroke))
colnames(rescount)[2] <- 'Stroke'

# Bar Chart of Patients base on residency 
ggplot(rescount, aes(x = Var1, y = Freq, fill = Stroke)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Freq), vjust = 0) +
  labs(title="Residence Type of the Patients",x ="Residence Type", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5))
```

There is virtually an even distribution between records in Rural residence and urban ones. Moreover,there is no significat effect of residency on stroke_experienced number of patients.

**9.**

```{r}
# Create smoking status counts table
smokecount= as.data.frame(table(heart_data$smoking_status,heart_data$stroke))
colnames(smokecount)[2] <- 'Stroke'
# Bar Chart of Patients based on smoking status
ggplot(smokecount, aes(x = Var1, y = Freq, fill = Stroke)) +
  geom_bar(stat = "identity")+
  geom_text(aes(label = Freq), vjust = 1.5) +
  labs(title="Smoking Status of Patients",x ="Smoking Status", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5))
```

The majority of data have never smoked.the data for formerly and currently smokers are similar. As illustrated,surprisingly,there is no significant effect on people smoking status and stroke.

### Histograms

The second tool we use is Histogram charts to illustrates the relationship between Quantifiable features: bmi,age and average glucose level.

**1.**

```{r}
#Histogram of Age with normal distribution overlay
histage=hist(heart_data$age,xlim=c(0,95),
                main="Histogram of Age with Normal Distribution Overlay",
                xlab="Age",las=1,col="lightblue")
xfit=seq(min(heart_data$age),max(heart_data$age))
yfit=dnorm(xfit,mean=mean(heart_data$age),sd=sd(heart_data$age))
#scaling
yfit=yfit*diff(histage$mids[1:2])*length(heart_data$age)
lines(xfit,yfit,col="red",lwd=2)
mean(heart_data$age)
median(heart_data$age)
```

As illustrated, the age of patients in this study is close to normal distribution of age which got the mean of **43.22661** and based on this information we can conclude the majority of record are around **40** years old.

**2.**

```{r}
#Histogram of Average Glucose Level with normal distribution overlay
histglucose=hist(heart_data$avg_glucose_level,xlim=c(0,300),
                    main="Histogram of Avg. Glucose with Normal Distribution Overlay",
                    xlab="Avg. Glucose",las=1,col="pink")
xfit =seq(min(heart_data$avg_glucose_level),max(heart_data$avg_glucose_level))
yfit =dnorm(xfit,mean=mean(heart_data$avg_glucose_level),sd=sd(heart_data$avg_glucose_level))
#scaling
yfit = yfit*diff(histglucose$mids[1:2])*length(heart_data$avg_glucose_level)
lines(xfit,yfit,col="red",lwd=2)
mean(heart_data$avg_glucose_level)
median(heart_data$avg_glucose_level)
```

The average glucose levels of the patients in the study are **right skewed**,with mean of **106.1477** which is greater than median of **91.885**.

### Box Plot

The third tool we use is Box plot for comparing one numeric and one quality features such as stroke and average glucose level and BMI.

**1.**

```{r}
# Box and Whisker Plot of Average Glucose in Patients With and Without Strokes
boxplot(Yes$avg_glucose_level,No$avg_glucose_level,
        main="Boxplot of Average Glucose Level by Stroke Status",
        ylab="Average Glucose Level",las=1,names=c("Stroke","No Stroke"))
```

The boxplot shows a relatively similar mean average glucose level in patients who suffered strokes and patients who have not, with lots of high outliers among non-stroke victims. we can conclude the average glucose level is not an effective feature on stroke.

**2.**

```{r}
#Box and Whisker Plot of Body Mass Index(BMI)in Patients With and Without Strokes
boxplot(Yes$bmi,No$bmi,main="Boxplot of Body Mass Index by Stroke Status",
        ylab="BMI",las=1,names=c("Stroke","No Stroke"))
```

The mean of both patient with stroke and without it is relatively similar, we can see the BMI is not the most effective feature on stroke.

**3.**

```{r}
boxplot(Yes$age,No$age,main="Boxplot of Age by Stroke Status",
        ylab="age",las=1,names=c("Stroke","No Stroke"))
```

As it is shown, the older the patient is, the mean of stoke is higher than no stoke, so we can say age is a significant factor on stroke.

### **Point Plot**

And as for the last tool, I used the ggplot2 library to draw a point chart for relation ship between age and stroke chance.

```{r}
ggplot(data=heart_data,aes(age,stroke))+
  geom_jitter(aes(color=factor(stroke)),alpha=0.7)+theme_light()
```

As the Barcharts above illustrated, the majority of patients did not have stoke, but among those who had they are mostly from the age 40 to older.(mostly between 60 to 80.

## Processing Data

As we saw in above visualizations, we have several not significant variables: "gender","residence type","bmi","woke type,"id". So I decided to remove them from the main data. Also to prevent further confusion between the positive and vegetative class, I change stroke=1 to Y and stroke=0 to N.

```{r}
heart_data=subset(heart_data,select=-c(id,gender,Residence_type,bmi,work_type))
heart_data$stroke=ifelse(heart_data$stroke=="1","Y","N")
heart_data$stroke=factor(heart_data$stroke)
```

Now we continue with splitting the "heart_data" into "Train" and "Test" . and the "train" data to "Estimation" and "Validation".

```{r}
#split Train and test

set.seed(42)
heart_trn_idx = sample(nrow(heart_data), size = 0.8 * nrow(heart_data))
heart_trn = heart_data[heart_trn_idx, ]
heart_tst = heart_data[-heart_trn_idx, ]

#split estimation and validation 
heart_est_idx = sample(nrow(heart_trn), size = 0.8 * nrow(heart_trn))
heart_est = heart_trn[heart_est_idx, ]
heart_val = heart_trn[-heart_est_idx, ]
#checking data
levels(heart_data$stroke)
prop.table(table(heart_data$stroke))
GGally::ggpairs(heart_data, progress = FALSE)

```

## Criteria Function

In this question ,it is important for us to predict the positive(Y) factors correctly, since it's important to knon whether a person is likely to have **stroke** or not in order to be able to **prevent** it. so the criteria we use is based on the model with **least False positiv**e because we do not want the patient **likely** to have stroke be left with **no alert** with wrong prediction.

there fore we I created a criteria function to calculate only Sensitivity and Accuracy(two important criteria).

```{r}
calc_criteria = function(pred1) {
  
  TP = sum(heart_val$stroke == "Y" & pred1 == "Y")
  TN = sum(heart_val$stroke== "N" & pred1 == "N")
  FP = sum(heart_val$stroke == "N" & pred1 == "Y")
  FN = sum(heart_val$stroke == "Y" & pred1 == "N")
  P = TP + FN
  N = TN + FP
  
  c(acc = (TP + TN) / (P+N),
    sens = TP/P)
  
}
```

## None Parametric Methods

### KNN

```{r}
heart_knn_est=heart_est
heart_knn_val=heart_val
```

#### normalization

Before we start with fitting candidate models of KNN we have to use Scale function to normalize the quantifiable features such as age and glucose level.

```{r}
#age
heart_knn_est$age.s=scale(heart_knn_est$age)
age.center = attr(heart_knn_est$age.s,"scaled:center")
age.scale = attr(heart_knn_est$age.s,"scaled:scale")
#average glucose level
heart_knn_est$avg_glucose_level.s=scale(heart_knn_est$avg_glucose_level)
avg_glucose_level.center = attr(heart_knn_est$avg_glucose_level.s,"scaled:center")
avg_glucose_level.scale = attr(heart_knn_est$avg_glucose_level.s,"scaled:scale")

# normalization of validation data
#age
heart_knn_val$age.s = scale(heart_knn_val$age, center = age.center, scale = age.scale)
#average glucose level
heart_knn_val$avg_glucose_level.s = scale(heart_knn_val$avg_glucose_level, center = avg_glucose_level.center, scale = avg_glucose_level.scale)
```

#### Tuning K

Now after we normalized our features, I created a function to generate K ranging from 1 to 100 and after comparing the Sensitivity and Accuracy we fit the final model.

```{r}

set.seed(42)
knn_mod_list=list()
creat_knn_normal=function(i){
  knn3(stroke~ hypertension+age.s+avg_glucose_level.s+heart_disease+ever_married+smoking_status, data = heart_knn_est, k =i)
}
for(i in 1:20){
  knn_mod_list[i]=list(creat_knn_normal(i))
}
#length(knn_mod_list)
#validation prediction
knn_val_pred=lapply(knn_mod_list,predict,heart_knn_val,type="class")
knn_val_mis=sapply(knn_val_pred,calc_criteria)
kable(as.data.frame(knn_val_mis))
# plot results
barplot(knn_val_mis,beside = T,legend.text = T,col=c("lightpink","green"))
```

As the result shows, the KNN with k=1 with the 90% accuracy and 12% sensitivity is the candid model.

#### Regression Tree

In this method I try to fit suitable models for predicting stroke occurring, I use the Regression Tree to split the data based on important features and to achieve a final node. With regression tree ,I create a list with 10 models, 5 consist of minsplit =5 and other 20 with cp ranging from 0 to 1.

```{r}
#tree
tree_mod_list = list(
  tree_1 = rpart(stroke~ ., data = heart_est, cp = 0.000,minsplit=5),
  tree_2 = rpart(stroke~ ., data = heart_est, cp = 0.001,minsplit=5),
  tree_3 = rpart(stroke~ ., data = heart_est, cp = 0.01,minsplit=5),
  tree_4 = rpart(stroke~ ., data = heart_est, cp = 0.1,minsplit=5),
  tree_5 = rpart(stroke~ ., data = heart_est, cp = 1,minsplit=5),
  tree_6 = rpart(stroke~ ., data = heart_est, cp = 0.000,minsplit=20),
  tree_7 = rpart(stroke~ ., data = heart_est, cp = 0.001,minsplit=20),
  tree_8 = rpart(stroke~ ., data = heart_est, cp = 0.01,minsplit=20),
  tree_9 = rpart(stroke~ ., data = heart_est, cp = 0.1,minsplit=20),
  tree_10 =rpart(stroke~ ., data = heart_est, cp = 1,minsplit=20)
)

tree_val_pred=lapply(tree_mod_list,predict,heart_val,type="class")
tree_val_mis=sapply(tree_val_pred,calc_criteria)
barplot(tree_val_mis,beside = T,legend.text = T,col=c("purple","yellow"))
rpart.plot(tree_mod_list$tree_2)
```

As illustration shows, the model 2 with accuracy of 91% which is slighlty higher than model 1 and sensitivity of 8% is the candid one.

## Parametric Methods

### Logistic Regression

Before We start to fit the models and check different cutoffs, we should know that Stroke=Y is positive for us because it's the **important factor** in this problem.

First we use the Function below to be able to compare different cutoffs criteria of each model.

```{r}
set.seed(42)
calc_metrics_cutoff = function(probs, cutoff) {
  
  pred1 = factor(ifelse(probs > cutoff, "Y", "N"))
  
  TP = sum(heart_val$stroke == "Y" & pred1 == "Y")
  TN = sum(heart_val$stroke== "N" & pred1 == "N")
  FP = sum(heart_val$stroke == "N" & pred1 == "Y")
  FN = sum(heart_val$stroke == "Y" & pred1 == "N")
  P = TP + FN
  N = TN + FP
  Ps = TP + FP
  Ns = TN + FN
  
  c(acc = (TP + TN) / (P+N),
    sens = TP/P,
    spec = TN/N)
  
}
```

Now, we fit different logistic with different cutoff models.

```{r}
# trying cutoffs
#model 1
reg_mod1= glm(stroke ~ ., data = heart_est, family = binomial)
reg_probs1 = predict(reg_mod1, heart_val, type = "response")
cutoffs = seq(from = 0.1, to = 1, by = 0.2)
results1 = sapply(cutoffs, calc_metrics_cutoff, probs = reg_probs1)
results1 = as.data.frame(t(rbind(cutoffs, results1)))
kable(results1)
pred1 = factor(ifelse(reg_probs1 > 0.1, "Y", "N"))
CM1 = table(pred1,heart_val$stroke)


#model 2
reg_mod2 =glm(stroke ~ age+avg_glucose_level, data = heart_est, family = binomial)
reg_probs2 = predict(reg_mod2, heart_val, type = "response")
cutoffs = seq(from = 0.1, to = 1, by = 0.2)
results2 = sapply(cutoffs, calc_metrics_cutoff, probs = reg_probs2)
results2 = as.data.frame(t(rbind(cutoffs, results2)))
kable(results2)
pred2 = factor(ifelse(reg_probs2 > 0.1, "Y", "N"))
CM2 = table(pred2,heart_val$stroke)

#model 3
reg_mod3 =glm(stroke ~ age+avg_glucose_level+hypertension+heart_disease, data = heart_est ,family = binomial)
reg_probs3 = predict(reg_mod3, heart_val, type = "response")
cutoffs = seq(from = 0.1, to = 1, by = 0.2)
results3 = sapply(cutoffs, calc_metrics_cutoff, probs = reg_probs3)
results3 = as.data.frame(t(rbind(cutoffs, results3)))
kable(results3)
pred3 = factor(ifelse(reg_probs3 > 0.1, "Y", "N"))
CM3 = table(pred3,heart_val$stroke)

#model 4
reg_mod4 =step(glm(stroke ~ .^2, data = heart_est ,family = binomial),trace=FALSE)
reg_probs4 = predict(reg_mod4, heart_val, type = "response")
cutoffs = seq(from = 0.1, to = 1, by = 0.2)
results4 = sapply(cutoffs, calc_metrics_cutoff, probs = reg_probs4)
results4 = as.data.frame(t(rbind(cutoffs, results4)))
kable(results4)
pred4 = factor(ifelse(reg_probs4 > 0.1, "Y", "N"))
CM4 = table(pred4,heart_val$stroke)
```

As a result of these 4 confusion matrix illustrates base on highest accuracy, the fourth model consist of all features with their binary interaction has the most Accuracy in cutoff 0.5. this cut off also has the sensitivity of 0.04 which means the number of True positives(stroke=1, pred=1) are not so much high. out of all the 50 positive(stroke=1), we only have predicted 2 of them right.(1-48/50)=0.04! on the other hand , the important factor, specificity is high. specificity indicates on (stroke=0 and pred=0) TN which is 768 out of 768, so the specificity is 1 and it means that we make no mistake in prediction of people unlikely to have stroke. accuracy is 94% which means that model is doing Alright specially predicting the healthy patients being healthy.What matters most in this modeling is that we should not have high FN. which means we do not say a patients who is likely to have stroke that they are healthy. in this model out of 50 people likely to have stoke, we wrongly predicted 49 of them that they are not going to have a stroke, so the model is not good at all.

So we should switch from the criteria of choosing the models based on the accuracy to highest **Sensitivity.**

```{r}
log_table=cbind(
  calc_criteria(pred1),
  calc_criteria(pred2),
  calc_criteria(pred3),
  calc_criteria(pred4)
)
barplot(log_table,beside = T,col=c("lightblue","lightpink"),legend.text = T)
```

As we can see the models with most sensitivity is model 2 and 3 with 0.54%. how ever our candid Logistic model is model 3, since the second important factor , accuracy is slightly higher.

## Discrimination models

In this Part, I first organized a list of LDA , QDA and Naive Bayes models which I thought can work efficaciously , and then I put prediction of them in another list to be able to set the important criteria on them and compare them to determine the candid model of each method.

### LDA

```{r}
#LDA
lda_mod1=lda(stroke~. , data=heart_est)
lda_mod2=lda(stroke~poly(age+avg_glucose_level ,degree=2),data=heart_est)
lda_mod3=lda(stroke~ . , data=heart_est,prior=c(1,1)/2)
lda_mod4=lda(stroke~hypertension+heart_disease+age+avg_glucose_level,data=heart_est)
lda_list=list(
  pred_lda1=predict(lda_mod1,heart_val)$class,
  pred_lda2=predict(lda_mod2,heart_val)$class,
  pred_lda3=predict(lda_mod3,heart_val)$class,
  pred_lda4=predict(lda_mod4,heart_val)$class
  
)
attach(lda_list)
criteria_val=sapply(lda_list,calc_criteria)

table_lda=rbind(calc_criteria(pred_lda1),
      calc_criteria(pred_lda2),
      calc_criteria(pred_lda3),
      calc_criteria(pred_lda4)
)
barplot(t(table_lda),besid=T,legend.text = T,col=c("Yellow","Blue"))
```

As we can see the best model out of lda model is the one with uniform prior probabilities , since our main criteria here is not misclassifiaction or accuracy, we are looking for the model with highest sensitivity because it is important to predict right in order to raise survival rate of patients.but since the model is not perfect,we chose the third LDA model with uniform prior probabilities which had 77% accuracy and 68% sensitivity and doing better than other models in this criteria.

### QDA

```{r}
#QDA

qda_mod1=qda(stroke~. , data=heart_est)
qda_mod2=qda(stroke~poly(age+avg_glucose_level ,degree=2),data=heart_est)
qda_mod3=qda(stroke~ . , data=heart_est,prior=c(1,1)/2)
qda_mod4=qda(stroke~hypertension+heart_disease+age+avg_glucose_level,data=heart_est)
qda_list=list(
  pred_qda1=predict(qda_mod1,heart_val)$class,
  pred_qda2=predict(qda_mod2,heart_val)$class,
  pred_qda3=predict(qda_mod3,heart_val)$class,
  pred_qda4=predict(qda_mod4,heart_val)$class
  
)
attach(qda_list)
criteria_val=sapply(qda_list,calc_criteria)

table_qda=rbind(calc_criteria(pred_qda1),
                calc_criteria(pred_qda2),
                calc_criteria(pred_qda3),
                calc_criteria(pred_qda4)
)
barplot(t(table_qda),besid=T,legend.text = T,col=c("green","pink"))
```

Since our best model must have the highest sensitivity, The third model with accuracy of 76% and sensitivity of 70% is our candid model.

### Naive Bayes

```{r}
#Naive Bayes

nb_mod1=naiveBayes(stroke~. , data=heart_est)
nb_mod2=naiveBayes(stroke~ . , data=heart_est,prior=c(1,1)/2)
nb_mod3=naiveBayes(stroke~hypertension+heart_disease+age+avg_glucose_level,data=heart_est)
nb_list=list(
  pred_nb1=predict(nb_mod1,heart_val),
  pred_nb2=predict(nb_mod2,heart_val),
  pred_nb3=predict(nb_mod3,heart_val)
  
)
criteria_val=sapply(nb_list,calc_criteria)

pred_nb1=predict(nb_mod1,heart_val)
pred_nb2=predict(nb_mod2,heart_val)
pred_nb3=predict(nb_mod3,heart_val)
table_nb= rbind(calc_criteria(pred_nb1),
                calc_criteria(pred_nb2),
                calc_criteria(pred_nb3)
)
barplot(t(table_nb),besid=T,legend.text = T,col=c("purple","gold"))
```

The three models of Naive Bayes are all slightly different, the result shows that all of them have accuracy of nearly 87% and sensitivity around 35%. between three, the third model is with the most accuracy and sensitivity, so we pick the third model in this method.

## Visualization and Testing

### Comparing all

```{r}
#creating a list of candid models of all methods

Candid_list=list(
  KNN=knn3(stroke~ hypertension+age.s+avg_glucose_level.s+heart_disease+ever_married+smoking_status, data = heart_knn_est, k =1),
  Tree=rpart(stroke~ ., data = heart_est, cp = 0.001,minsplit=5),
  GLM=glm(stroke ~ age+avg_glucose_level+hypertension+heart_disease, data = heart_est ,family = binomial),
  LDA=lda(stroke~ . , data=heart_est,prior=c(1,1)/2),
  QDA=qda(stroke~ . , data=heart_est,prior=c(1,1)/2),
  NB=naiveBayes(stroke~hypertension+heart_disease+age+avg_glucose_level,data=heart_est)
)
KNN=knn3(stroke~ hypertension+age.s+avg_glucose_level.s+heart_disease+ever_married+smoking_status, data = heart_knn_est, k =1)
Tree=rpart(stroke~ ., data = heart_est, cp = 0.001,minsplit=5)
GLM=glm(stroke ~ age+avg_glucose_level+hypertension+heart_disease, data = heart_est ,family = binomial)
LDA=lda(stroke~ . , data=heart_est,prior=c(1,1)/2)
QDA=qda(stroke~ . , data=heart_est,prior=c(1,1)/2)
NB=naiveBayes(stroke~hypertension+heart_disease+age+avg_glucose_level,data=heart_est)
Pred_list=list(
  pred_knn=predict(KNN,heart_knn_val,type="class"),
  pred_tree=predict(Tree,heart_val,type="class"),
  pred_glm=factor(ifelse(predict(GLM, heart_val, type = "response") > 0.1, "Y", "N")),
  pred_lda=predict(LDA,heart_val)$class,
  pred_qda=predict(QDA,heart_val)$class,
  pred_nb=predict(NB,heart_val)
)
all_sens_val=sapply(Pred_list,calc_criteria)
kable(all_sens_val)
```

### Visualization

```{r}
#visualization
x=c("KNN","Tree","GLM","LDA","QDA","NB")
table1=as.data.frame(cbind(x,all_sens_val[1,]))
plot1=ggplot(data=table1,aes(x=x,y=V2,group=1))+geom_line(color="purple",size=1)+geom_point(col="red",size=3)+theme_light()+
  xlab("Models")+ylab("Accuracy")
table2=as.data.frame(cbind(x,all_sens_val[2,]))
plot2=ggplot(data=table2,aes(x=x,y=V2,group=1))+geom_line(color="lightblue",size=1)+geom_point(col="darkgreen",size=3)+theme_light()+
  xlab("Models")+ylab("sensitivity")
plot3=plot_grid(plot1,plot2)
title3=ggdraw()+draw_label("Accuracy and Sensitivity of each Model")
plot4=plot_grid(title3,plot3,ncol=1,rel_heights=c(0.1, 1))
plot4
```

As both result of tables and chart illustrates, the highest sensitivity belongs to model LDA andQDa , since the accuracy of model QDA is 1% higher than LDA , we chose QDA model as our final Candid Model.

### Testing

Now we fit the final Candid model, QDA, on train data and then test the prediction on test data.

```{r}
#fitting the final model on train
Final_mod=qda(stroke~ . , data=heart_trn,prior=c(1,1)/2)
#predict on test
Final_pred=predict(Final_mod,heart_tst,type="response")$class
confusionMatrix(as.factor(Final_pred), heart_tst$stroke, positive = "Y")
table(Final_pred,heart_tst$stroke)

```

We can see the final Model has the sensitivity of 70% and accuracy of 74% , in general the model is doing alright.

## Over and Under Sampling

### Over

I guess the problem of not accurately predicting the patients with stroke is due to this problem so we use Over Sampling.

```{r}
#Over Sampling

set.seed(42)

over_data = ovun.sample(stroke~ . , data=heart_trn, method = "over")$data

table(over_data$stroke)

qda.fit.over = qda( stroke ~ . , data = over_data,prior=c(1,1)/2, family = binomial)

pred.over =predict(qda.fit.over, heart_tst , type = "response")$class

confusionMatrix(as.factor(pred.over), heart_tst$stroke, positive = "Y")


```

With this change, we can see that specificity and accuracy did not change significantly,sensitivity raised up about 2%.

### under

```{r}
#Under Sampling
under_data = ovun.sample(stroke~ . , data=heart_trn, method = "under")$data
table(under_data$stroke)

qda.fit.under = qda( stroke ~ . , data = under_data,prior=c(1,1)/2, family = binomial)

pred.under =predict(qda.fit.under, heart_tst , type = "response")$class

confusionMatrix(as.factor(pred.under), heart_tst$stroke, positive = "Y")
```

With this change, we can see that specificity and accuracy did not change significantly , but the sensitvity went up 2%.

## Conclusion

This Report is created by Rmarkdown and aims to illustrate the analysis and reasons behind the chunks of R codes of Machine learningClassifiaction project on stroke occurring prediction and statistical calculations.

I hope you a have found this Report appropriate to your use.

Thank you for your attention.

*Mina Kanaani.*

 
